### 2019-01-21 writing/thesis/chapters/results.tex
Include corrections from proof reader.

Before computing the error we first need to define the $L_p$ norms for a $p > 0$ by
% notation: https://en.wikipedia.org/wiki/Lp_space#Lp_spaces
\begin{equation}
  \label{eq:Lp-nrom}
  \Vert f(x) \Vert_p = \left( \intddomain{ \vert f(\bm{x}) \vert^p}  \right)^{1/p}.
\end{equation}
We are interested in the total error which we define as the norm of the difference between an analytical solution $f(\bm{x}, t)$ and our approximation $\hat{f}(\bm{x}, t)$.
The error at a time $t$ is thus defined as
\begin{equation}
  \label{eq:error}
  \error(t,p) = \Vert f(\bm{x}, t) - \hat{f}(\bm{x}, t) \Vert_p.
\end{equation}

We first observe that for our approximation space $\broken$ we can split up the error as a sum over all cells.
We evaluate the cell-wise error with Gaussian quadrature~\pcref{eq:quadrature,eq:integration-by-substitution}
 \begin{equation}
%   \Vert f(\bm{x}, t) - \hat{f}(\bm{x}, t) \Vert_p^p =
   \error(t,p) = 
   \left( \sum_{\cell[i] \in \broken}
    \left( V \sum_{\bm{j}} \vert f^{\cell[i]}(\quadNode[j], t) - \hat{f}^{\cell{i}}(\quadNode[j], t) \vert^p \, \quadWeight[j] \right) \right)^{1/p}.
 \end{equation}
Here, a superscript of $\cell[j]$ denotes the restriction of the function to a grid cell.
To increase the quadrature quality, we evaluate both analytical and approximate solution at the quadrature nodes of a scheme with 10 quadrature nodes $\quadNode[j]$ with weights $\quadWeight[j]$.
We then use these nodes to compute the approximation error per cell.

@@

\begin{figure}[htb]
  \centering
  \includegraphics[trim=0.2cm 0.3cm 0.2cm 1.0cm, clip]{thesis_convergence}
  \caption{$L_2$-Error vs.\ Grid Size.
    The dashed lines are the optimal order of convergence $N+1$.
  The line is estimated by a linear regression fit and and setting the slope to the ideal order.}
  \label{fig:convergence-l2-error}
\end{figure}

@@

\sidetitle{Lid-Driven Cavity}
\begin{figure}[htb]
  \centering
  \begin{subfigure}[t]{0.5\textwidth}
    \centering
    \includegraphics[trim={0.05cm 0.05cm 0.55cm 1.0cm},clip]{thesis_lid_driven_cavity_stream}
    \caption{\label{fig:lid-driven-cavity-streamlines}%
      Streamlines of the cavity flow.}
  \end{subfigure}~%
  \begin{subfigure}[t]{0.5\textwidth}
    \centering
    \includegraphics{thesis_lid_driven_cavity}
    \caption{\label{fig:lid-driven-cavity-result}%
      Plot of velocity slices. Lines are our soulution, scatter plot shows reference solution~\cite{ghia1982high}.
      For each slice, the other variable is hold constant at a value of zero.}
  \end{subfigure}

@@

The coarse structure of our results~(\cref{fig:cosine-bubble-contour-ours}) agrees with the reference solution~(\cref{fig:cosine-bubble-contour-giraldo} of~\cite{giraldo2008study} but the fine details differ.
In contrast to us, they used a scheme of polynomial order 10 and a uniform spatial resolution of \SI{20}{\m}.
Furthermore, we used a physical viscosity of $0.1$, which explains why we have fewer spurious details while using a lower order.
This becomes clearer by comparing the potential temperature pertubation.
Our solution has a smooth distribution inside the cloud~(\cref{fig:cosine-bubble-potT-ours}) while the results of~\cite{giraldo2008study} show sharp peaks~(\cref{fig:cosine-bubble-potT-giraldo}).
In addition, our solution does not contain negative values for the pertubation.
Despite the different parameters, the maximum pertubation and its position matches for both results.

@@

\section{AMR vs Static Mesh}\label{sec:results-tts-amr}
To show the effectiveness of the \amr{} criterion, we compare the two settings:
\begin{description}
\item[AMR] We used a coarse mesh with just $27 \times 27$ cells.
  Each could can be refined once.
  That is, the cells had sidelengths of roughly \SI{37.04}{\m} and  \SI{12.35}{\m}.
  As refinement thresholds we used $T_\text{refine} = 2.5$ and $T_\text{delete} = -0.5$ in \cref{eq:refinement-criterion}.
  They are the exact same settings that we used for the bubble scenarios in the previous section.
\item[Fine] We used a uniform grid where all cells had a sidelength corresponding to the finest mesh of the \amr{}-test case.
  In addition, we disabled the reduction of global observables.
\end{description}
Note, that while our \amr{} criterion is able to cope with an additional, coarser level, the current implementation in \exahype{} is unstable in this case.
This is why we only report an \amr{} run with one level of dynamic refinement.

% \begin{figure}[htb]
%   \centering
%   \includegraphics{thesis_two_bubbles_contour_t600_noamr} 
%   \caption{\label{fig:two-bubbles-noamr-contour}%
%     Two bubbles run with fully refined mesh}
% \end{figure}
For both settings, we used an \aderdg{}-method of order $4$ and thus have a minimal effective resolution of ca.\ $\SI{3.09}{\m}$.
%The final result for the run with the fully refined mesh (\cref{fig:two-bubbles-noamr-contour}) looks identical to the \amr{} results (\cref{fig:two-bubbles-contour}) shown in the previous section.
Note that the \amr{} criterion manages to track the bubble correctly~(\cref{fig:two-bubbles-amr-grid}) and introduces only small errors~(\cref{fig:two-bubbles-amr-error}) compared to the solution with the fully refined mesh.

@@

  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \includegraphics{thesis_detonation_euler_nodmp_Z}
    \caption{$Z$, Euler, no \textsc{dmp}}
  \end{subfigure}~%
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \includegraphics{thesis_detonation_euler_dmp_Z}
    \caption{$Z$, Euler, \textsc{dmp}}
  \end{subfigure}~%
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \includegraphics{thesis_detonation_ns_Z}
    \caption{$Z$, \textsc{NS}, no \textsc{dmp}}
  \end{subfigure}
  \caption{\label{fig:detonation-line-plots}%
    Slices along the $x$-axis for our three configurations.
    The coordinate $z = 0$ is held constant.
    We only show one half of the slice because the scenario is symmetric.
  }
\end{figure}
The results for runs with and without \textsc{dmp} can be seen in \cref{fig:detonation-line-plots}.
Note, that oscillations appeared for the Euler equations unless we use the \textsc{dmp}-criterion.
This is not the case for the Navier Stokes equations.
This scenario shows the differences between the viscous Navier-Stokes equations and the Euler equations clearly:
The viscous effects smooth out the otherwise sharp curves.

Furthermore, they have the effect that fewer cells are limited~(\cref{fig:detonation-pressure-contour}).
Note that the limiter activation are not symmetric which is unexpected for a radial-symmetric scenario.
The contours for pressure and mass fraction look similar for all considered settings.

It is difficult to compare these results with our references, because they both use different settings.
In constrast to~\cite{helzel2000modified} we use a non-stiff source term and viscous effects,~\cite{hidalgo2011ader} includes viscosity but is limited to one-dimensional scenarios.
Comparing with \cref{fig:detonation-contour-reference} we can see that our simulation captures the overall behaviour of the reaction similar to the stiff simulations of~\cite{helzel2000modified}.
The viscous effects have a similar smoothing effect as in~\cite{hidalgo2011ader}.

>>>>

Before computing the error, we first need to define the $L_p$ norms for a $p > 0$ by
\begin{equation}
  \label{eq:Lp-nrom}
  \Vert f(\bm{x}) \Vert_p = \left( \intddomain{ \vert f(\bm{x}) \vert^p}  \right)^{1/p}.
\end{equation}
We are interested in the total error which we define as the norm of the difference between an analytical solution $f(\bm{x}, t)$ and our approximation $\hat{f}(\bm{x}, t)$.
The error at a time $t$ is thus defined as
\begin{equation}
  \label{eq:error}
  \error(t,p) = \Vert f(\bm{x}, t) - \hat{f}(\bm{x}, t) \Vert_p.
\end{equation}

We first observe that for our approximation space $\broken$ we can split up the error as a sum over all cells.
We evaluate the cell-wise error with Gaussian quadrature~\pcref{eq:quadrature,eq:integration-by-substitution}
 \begin{equation}
%   \Vert f(\bm{x}, t) - \hat{f}(\bm{x}, t) \Vert_p^p =
   \error(t,p) = 
   \left( \sum_{\cell[i] \in \broken}
    \left( V \sum_{\bm{j}} \vert f^{\cell[i]}(\quadNodeNd[j], t) - \hat{f}^{\cell{i}}(\quadNodeNd[j], t) \vert^p \, \quadWeight[j] \right) \right)^{1/p}.
 \end{equation}
Here, a superscript of $\cell[j]$ denotes the restriction of the function to a grid cell.
To increase the quadrature quality, we evaluate both analytical and approximate solution at the quadrature nodes of a scheme with 10 quadrature nodes $\quadNode[j]$ with weights $\quadWeight[j]$.
We then use these nodes to compute the approximation error per cell.

@@

\begin{figure}[htb]
  \centering
  \includegraphics[trim=0.2cm 0.3cm 0.2cm 1.0cm, clip]{thesis_convergence}
  \caption{$L_2$-Error vs.\ Grid Size.
    The dashed lines are the optimal order of convergence $N+1$.
  The line is estimated by a linear regression fit and setting the slope to the ideal order.}
  \label{fig:convergence-l2-error}
\end{figure}

@@

\sidetitle{Lid-Driven Cavity}
\begin{figure}[htb]
  \centering
  \begin{subfigure}[t]{0.5\textwidth}
    \centering
    \includegraphics[trim={0.05cm 0.05cm 0.55cm 1.0cm},clip]{thesis_lid_driven_cavity_stream}
    \caption{\label{fig:lid-driven-cavity-streamlines}%
      Streamlines of the cavity flow.}
  \end{subfigure}~%
  \begin{subfigure}[t]{0.5\textwidth}
    \centering
    \includegraphics{thesis_lid_driven_cavity}
    \caption{\label{fig:lid-driven-cavity-result}%
      Plot of velocity slices. Lines are our soulution, scatter plot shows reference solution~\cite{ghia1982high}.
      For each slice, the other variable is held constant at a value of zero.}
  \end{subfigure}

@@

The coarse structure of our results~(\cref{fig:cosine-bubble-contour-ours}) agrees with the reference solution~(\cref{fig:cosine-bubble-contour-giraldo} of~\cite{giraldo2008study} but the fine details differ.
In contrast to us, they used a scheme of polynomial order 10 and a uniform spatial resolution of \SI{20}{\m}.
Moreover, we used a physical viscosity of $0.1$, which explains why we have fewer spurious details despite using a lower order.
This becomes clearer by comparing the potential temperature pertubation.
Our solution has a smooth distribution inside the cloud~(\cref{fig:cosine-bubble-potT-ours}) while the results of~\cite{giraldo2008study} show sharp peaks~(\cref{fig:cosine-bubble-potT-giraldo}).
In addition, our solution does not contain negative values for the pertubation.
Despite the different parameters, the maximum pertubation and its position matches for both results.

@@

\section{AMR vs Static Mesh}\label{sec:results-tts-amr}
To show the effectiveness of the \amr{} criterion, we compare the two settings:
\begin{description}
\item[AMR] We used a coarse mesh with just $27 \times 27$ cells.
  Each could can be refined once.
  That is, the cells had sidelengths of roughly \SI{37.04}{\m} and  \SI{12.35}{\m}.
  As refinement thresholds we used $T_\text{refine} = 2.5$ and $T_\text{delete} = -0.5$ in \cref{eq:refinement-criterion}.
  They are the exact same settings that we used for the bubble scenarios in the previous section.
\item[Fine] We used a uniform grid where all cells had a sidelength corresponding to the finest mesh of the \amr{}-test case.
  In addition, we disabled the reduction of global observables.
\end{description}
While our \amr{} criterion is able to cope with an additional, coarser level, the current implementation in \exahype{} is unstable in this case.
This is why we only report an \amr{} run with one level of dynamic refinement.

% \begin{figure}[htb]
%   \centering
%   \includegraphics{thesis_two_bubbles_contour_t600_noamr} 
%   \caption{\label{fig:two-bubbles-noamr-contour}%
%     Two bubbles run with fully refined mesh}
% \end{figure}
For both settings, we used an \aderdg{}-method of order $4$ and thus had a minimal effective resolution of ca.\ $\SI{3.09}{\m}$.
%The final result for the run with the fully refined mesh (\cref{fig:two-bubbles-noamr-contour}) looks identical to the \amr{} results (\cref{fig:two-bubbles-contour}) shown in the previous section.
Note that the \amr{} criterion manages to track the bubble correctly~(\cref{fig:two-bubbles-amr-grid}) and introduces only small errors~(\cref{fig:two-bubbles-amr-error}) compared to the solution with the fully refined mesh.

@@

  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \includegraphics{thesis_detonation_euler_nodmp_Z}
    \caption{$Z$, Euler, no \textsc{dmp}}
  \end{subfigure}~%
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \includegraphics{thesis_detonation_euler_dmp_Z}
    \caption{$Z$, Euler, \textsc{dmp}}
  \end{subfigure}~%
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \includegraphics{thesis_detonation_ns_Z}
    \caption{$Z$, \textsc{NS}, no \textsc{dmp}}
  \end{subfigure}
  \caption{\label{fig:detonation-line-plots}%
    Slices along the $x$-axis for our three configurations.
    The coordinate $z = 0$ is held constant.
    We only show one half of the slice because the scenario is symmetric.
  }
\end{figure}
The results for runs with and without \textsc{dmp} can be seen in \cref{fig:detonation-line-plots}.
Note, that oscillations appeared for the Euler equations unless we use the \textsc{dmp}-criterion.
This is not the case for the Navier Stokes equations.

This scenario shows the differences between the viscous Navier-Stokes equations and the Euler equations clearly:
The viscous effects smooth out the otherwise sharp curves.
Furthermore, they have the effect that fewer cells are limited~(\cref{fig:detonation-pressure-contour}).
Note that the limiter activation are not symmetric which is unexpected for a radial-symmetric scenario.
The contours for pressure and mass fraction look similar for all considered settings.

It is difficult to compare these results with our references because they both use different settings.
In constrast to~\cite{helzel2000modified} we use a non-stiff source term and viscous effects; \cite{hidalgo2011ader} includes viscosity but is limited to one-dimensional scenarios.
Comparing with \cref{fig:detonation-contour-reference} we can see that our simulation captures the overall behaviour of the reaction similar to the stiff simulations of~\cite{helzel2000modified}.
The viscous effects have a similar smoothing effect as in~\cite{hidalgo2011ader}.

### 2019-01-21 writing/thesis/chapters/scenarios.tex
Cite manufactured solution notebook.
Include corrections from proof reader.

\section{A Manufactured Solution}\label{sec:manufactured-solution}
The compressible Navier Stokes equations have no interesting exact solution.
This is why we follow the method of manufactured solutions~\cite{salari2000code} and create our own.
Following~\cite{dumbser2010arbitrary}, we assume that
\begin{align}\label{eq:manufactured-solution}
\begin{split}
  \pressure(\bm{x}, t) &= \pressure_0 \cos( \bm{k} \bm{x} - \omega t ) + \pressure_b, \\
  \Qrho(\bm{x}, t) &= \Qrho_0 \sin (\bm{k} \bm{x} - \omega t) + \Qrho_b, \\
  \Qv(\bm{x}, t) &= \bm{v_0} \sin(\bm{k} \bm{x} - \omega t).
\end{split}
\end{align}
We use the constants
\begin{equation}
\begin{alignedat}{6}
  & \pressure_u &&= \nicefrac{1}{10}, \qquad& \pressure_b &= \nicefrac{1}{\gamma}, \\
  & \Qrho_0 &&= \nicefrac{1}{2}, \qquad& \Qrho_b &= 1, \\
  & \bm{v_0} &&= \nicefrac{1}{4},
  \begin{pmatrix}
    1 \\
    1
  \end{pmatrix},
  \qquad& \bm{k} &= \nicefrac{\pi}{5}
  \begin{pmatrix}
    1\\
    1
  \end{pmatrix},\\
  & \omega &&= 2 \pi.
  \end{alignedat}
\end{equation}
We derive a source term by inserting this solution into the \textsc{pde}~\pcref{eq:equation-set} by using the symbolic math toolkit \textit{Sage}~\cite{sagemath}.
Our computation of the source term can be found at XXX\todo{Cite github repository with notebook!}.
In combination with the derived source term, \cref{eq:manufactured-solution} then is an exact solution of our \pde{}.

@@

\sidetitle{CJ Detonation Wave}
We investigate the radialsymmetric Chapman-Jouget (\textsc{cj}) detonation wave scenario from~\cite{helzel2000modified}.
This scenario for the reactive Navier Stokes equations is set up in a 2D box of size $[-1, 1] \times [-1, 1]$ surrounded by no-slip walls~\pcref{eq:no-slip}.
We use the reactive source term~\pcref{eq:source-reaction,eq:reaction-model} and set the characteristic timescale to the non-stiff case $\reactionTimescale = 0.1$ and the threshold temperature to $\reactionTemperature = 0.26$.
Inside a circle of radius $0.3$ we have totally unburnt gas, outside of this circle we have burned gas.
Let $\alpha =  \operatorname{atan2}(z,x)$ be the angle in polar coordinates.
We use the initial conditions
\begin{equation}\label{eq:cj-detonation}
\begin{alignedat}{6}
  & \Qrho_u &&= 0.887565 \qquad& \Qrho_b &= 1.4\\
  & \Qv_u &&= -0.577350
  \begin{pmatrix}
     \cos (\alpha)\\
     \sin (\alpha)
   \end{pmatrix}  \qquad& \Qv_b  &=
   \begin{pmatrix}
     0 \\
     0
   \end{pmatrix}\\
   & \pressure_u &&=0.191709 \qquad& \pressure_b &= 1\\
   & \QZZ_u &&= 1 \quad& \QZZ_b &= 0
  \end{alignedat}
\end{equation}
where a subscript of $u$ and $b$ denotes the unburnt and burnt state respectively.
Note that this scenario is radialsymmetric.

@@

The initial conditions for pressure and density can be computed directly from the \textsc{pde} once few assumptions have been given.
We assume that $\backgroundPotT$ is constant over the entire domain~\cite{giraldo2008study}.
Inserting the definition of potential temperature, given by \cref{eq:potTToT}, into the equation of state~\pcref{eq:eos} leads us to
\begin{equation}
  \pressure(z) = \Qrho(z) R {\left( \backgroundPotT \frac{p}{p_0}\right)}^{R/c_p},
\end{equation}
which we solve for $p(z)$.
After some algebra and simplifying, we arrive at
\begin{equation}
 p(z) = p_0 \left( \frac{R \backgroundPotT \rho(z)}{p_0} \right)^{\gamma},
\end{equation}
or equivalently
\begin{equation}
  \label{eq:eos-potT}
\Qrho(z) = \frac{p_{0}^{\frac{R}{c_{p}}} p^{\frac{1}{\gamma}}{\left (z \right )}}{R \backgroundPotT}.
\end{equation}
We want to consider flow in hydrostatic equilibrium, i.e.\ flow where the force originating from the pressure-gradient is exactly balanced by gravitation.
Inserting this assumption in the momentum equation in $z$-direction and using \cref{eq:eos} once again, we arrive at the ordinary differential equation (\textsc{ode})
\begin{align}
  \label{eq:hydrostatic-balance-potT}
  \begin{split}
  \frac{d}{d z} p{\left (z \right )} &= -g \Qrho(z)
                                     = 
        - \frac{g p_{0}^{\frac{R}{c_{p}}} p^{\frac{1}{\gamma}}{\left (z \right )}}{R \backgroundPotT}\\
  p{\left (0 \right )} &= p_0
  \end{split}
\end{align}
where we use the reference pressure $p_0$ as the pressure on ground level.
This \textsc{ode} is the hydrostatic equilibrium (similar to \cref{eq:hydrostatic-balance}) in terms of the potential temperature and can be solved by seperation of variables.
After simplifying, we obtain
\begin{align}
  \label{eq:hydrostatic-background-pressure}
p(z) &= \left(\left(1 - \frac{1}{\gamma}\right) \left(C - \frac{g p_{0}^{\frac{R}{c_{p}}} z}{R \backgroundPotT}\right)\right)^{\frac{c_{p}}{R}},\\
\shortintertext{with constant of integration}
 C &= \frac{c_{p} p_{0}^{\frac{R}{c_{p}}}}{R}.
\end{align}

We can now compute the initial conditions for a scenario where the background potential energy is in hydrostatic equilbrium with a pertubation that is not in equilbrium.
We first compute the background pressure with background potential temperature $\backgroundPotT = \text{const.}$ using \cref{eq:hydrostatic-background-pressure}.
Next, we compute the pertubated potential temperature $\potT$ and use \cref{eq:potTToT} to convert it to temperature.
The density $\Qrho(\bm{x})$ can now be evaluated with the thermal equation of state \cref{eq:temperature}.
Finally, we compute the energy by inserting pressure and density into the equation of state~(\ref{eq:eos}).

@@

\sidetitle{Cosine Bubble}
As a simple test case we consider the cosine bubble scenario where we have one large warm air bubble in the middle of the domain~\cite{giraldo2008study}.
The initial conditions of the scenario are given by a pertubation of a hydrostatic equlibrium.
We use a background potential temperature of $\backgroundPotT = 300$ and a domain of $\SI{1}{\km} \times \SI{1}{\km}$.
Let
\begin{equation}
  \label{eq:radius}
  r^2 = (x - x_0)^2 + (z - z_0)^2
\end{equation}
denote the difference between spatial positions $x,z$ and the center of a bubble $x_c, z_c$.
The pertubation is given by the function
\begin{align}
  \label{eq:cos-pertubation}
  \pertubationPotT &= \begin{cases}
    \nicefrac{A}{2} \left[ 1 + \cos(\pi r) \right] & r \leq a, \\
    0 & r > a,
    \end{cases}
\end{align}
where $A$ denotes the maximal pertubation and $a$ is the size of the bubble.
We use the constants
\begin{equation}\label{eq:cosine-bubble}
  A = 0.5 \quad a = \SI{250}{\m} \quad x_c = \SI{500}{\m} \quad z_c = \SI{350}{\m}.
\end{equation}
We use a constant viscosity of $\mu = 0.1$.

\sidetitle{Two Bubbles}
As a simple test case we consider a scenario where we have one large warm air bubble in the middle and a small cold bubble on the top.
This scenario was first published by \citeauthor{robert1993bubble}~\cite{robert1993bubble}.
We use the same procedure as before to compute the initial conditions, again with $\backgroundPotT = 300$ but with a pertubation of the form

>>>>

\section{A Manufactured Solution}\label{sec:manufactured-solution}
The compressible Navier Stokes equations have no interesting exact solution.
This is why we follow the method of manufactured solutions~\cite{salari2000code} and create our own.
Following~\cite{dumbser2010arbitrary}, we assume that
\begin{align}\label{eq:manufactured-solution}
\begin{split}
  \pressure(\bm{x}, t) &= \pressure_0 \cos( \bm{k} \bm{x} - \omega t ) + \pressure_b, \\
  \Qrho(\bm{x}, t) &= \Qrho_0 \sin (\bm{k} \bm{x} - \omega t) + \Qrho_b, \\
  \Qv(\bm{x}, t) &= \bm{v_0} \sin(\bm{k} \bm{x} - \omega t).
\end{split}
\end{align}
We use the constants
\begin{equation}
\begin{alignedat}{6}
  & \pressure_u &&= \nicefrac{1}{10}, \qquad& \pressure_b &= \nicefrac{1}{\gamma}, \\
  & \Qrho_0 &&= \nicefrac{1}{2}, \qquad& \Qrho_b &= 1, \\
  & \bm{v_0} &&= \nicefrac{1}{4},
  \begin{pmatrix}
    1 \\
    1
  \end{pmatrix},
  \qquad& \bm{k} &= \nicefrac{\pi}{5}
  \begin{pmatrix}
    1\\
    1
  \end{pmatrix},\\
  & \omega &&= 2 \pi.
  \end{alignedat}
\end{equation}
We derive a source term by inserting this solution into the \textsc{pde}~\pcref{eq:equation-set} by using the symbolic math toolkit \textit{Sage}~\cite{sagemath}.
Our computation of the source term can be found at~\cite{krenz2018manufactured}.
In combination with the derived source term, \cref{eq:manufactured-solution} then is an exact solution of our \pde{}.

@@

\sidetitle{CJ Detonation Wave}
We investigate the radialsymmetric Chapman-Jouget (\textsc{cj}) detonation wave scenario from~\cite{helzel2000modified}.
This scenario for the reactive Navier Stokes equations is set up in a 2D box of size $[-1, 1] \times [-1, 1]$ surrounded by no-slip walls~\pcref{eq:no-slip}.
We use the reactive source term~\pcref{eq:source-reaction,eq:reaction-model} and set the characteristic timescale to the non-stiff case $\reactionTimescale = 0.1$ and the threshold temperature to $\reactionTemperature = 0.26$.
Inside a circle of radius $0.3$ we have completely unburnt gas, outside of this circle we have burned gas.
Let $\alpha =  \operatorname{atan2}(z,x)$ be the angle in polar coordinates.
We use the initial conditions
\begin{equation}\label{eq:cj-detonation}
\begin{alignedat}{6}
  & \Qrho_u &&= 0.887565 \qquad& \Qrho_b &= 1.4\\
  & \Qv_u &&= -0.577350
  \begin{pmatrix}
     \cos (\alpha)\\
     \sin (\alpha)
   \end{pmatrix}  \qquad& \Qv_b  &=
   \begin{pmatrix}
     0 \\
     0
   \end{pmatrix}\\
   & \pressure_u &&=0.191709 \qquad& \pressure_b &= 1\\
   & \QZZ_u &&= 1 \quad& \QZZ_b &= 0
  \end{alignedat}
\end{equation}
where a subscript of $u$ and $b$ denotes the unburnt and burnt state respectively.
Note that this scenario is radialsymmetric.

@@

The initial conditions for pressure and density can be computed directly from the \textsc{pde} once few assumptions have been given.
We assume that $\backgroundPotT$ is constant over the entire domain~\cite{giraldo2008study}.
Inserting the definition of potential temperature, given by \cref{eq:potTToT}, into the equation of state~\pcref{eq:eos} leads us to
\begin{equation}
  \pressure(z) = \Qrho(z) R {\left( \backgroundPotT \frac{p}{p_0}\right)}^{R/c_p},
\end{equation}
which we solve for $p(z)$.
After some algebra and simplifying, we arrive at
\begin{equation}
 p(z) = p_0 \left( \frac{R \backgroundPotT \rho(z)}{p_0} \right)^{\gamma},
\end{equation}
or equivalently
\begin{equation}
  \label{eq:eos-potT}
\Qrho(z) = \frac{p_{0}^{\frac{R}{c_{p}}} p^{\frac{1}{\gamma}}{\left (z \right )}}{R \backgroundPotT}.
\end{equation}
We want to consider flow in hydrostatic equilibrium, i.e.\ a flow for which the pressure-gradient and gravitational force are exactly in balance.
Inserting this assumption in the momentum equation in $z$-direction and using \cref{eq:eos} once again, we arrive at the ordinary differential equation (\textsc{ode})
\begin{align}
  \label{eq:hydrostatic-balance-potT}
  \begin{split}
  \frac{d}{d z} p{\left (z \right )} &= -g \Qrho(z)
                                     = 
        - \frac{g p_{0}^{\frac{R}{c_{p}}} p^{\frac{1}{\gamma}}{\left (z \right )}}{R \backgroundPotT}\\
  p{\left (0 \right )} &= p_0
  \end{split}
\end{align}
where we use the reference pressure $p_0$ as the pressure on ground level.
This \textsc{ode} is the hydrostatic equilibrium (similar to \cref{eq:hydrostatic-balance}) in terms of the potential temperature and can be solved by seperation of variables.
After simplifying, we obtain
\begin{align}
  \label{eq:hydrostatic-background-pressure}
p(z) &= \left(\left(1 - \frac{1}{\gamma}\right) \left(C - \frac{g p_{0}^{\frac{R}{c_{p}}} z}{R \backgroundPotT}\right)\right)^{\frac{c_{p}}{R}},\\
\shortintertext{with constant of integration}
 C &= \frac{c_{p} p_{0}^{\frac{R}{c_{p}}}}{R}.
\end{align}

We can now compute the initial conditions for a scenario with a background potential energy that is in hydrostatic equilbrium and a pertubation that is not.
We first compute the background pressure with background potential temperature $\backgroundPotT = \text{const.}$ using \cref{eq:hydrostatic-background-pressure}.
Next, we compute the pertubated potential temperature $\potT$ and use \cref{eq:potTToT} to convert it to temperature.
The density $\Qrho(\bm{x})$ can now be evaluated with the thermal equation of state \cref{eq:temperature}.
Finally, we compute the energy by inserting pressure and density into the equation of state~(\ref{eq:eos}).

@@

\sidetitle{Cosine Bubble}
As a simple test case, we consider the cosine bubble scenario that describes a background atmosphere that is pertubated by a large warm air bubble in the middle of the domain~\cite{giraldo2008study}.
We use a background potential temperature of $\backgroundPotT = 300$ and a domain of $\SI{1}{\km} \times \SI{1}{\km}$.
The background atmosphere is assumed to be in hydrostatic equilibrium.
Let
\begin{equation}
  \label{eq:radius}
  r^2 = (x - x_0)^2 + (z - z_0)^2
\end{equation}
denote the difference between spatial positions $x,z$ and the center of a bubble $x_c, z_c$.
The pertubation is given by the function
\begin{align}
  \label{eq:cos-pertubation}
  \pertubationPotT &= \begin{cases}
    \nicefrac{A}{2} \left[ 1 + \cos(\pi r) \right] & r \leq a, \\
    0 & r > a,
    \end{cases}
\end{align}
where $A$ denotes the maximal pertubation and $a$ is the size of the bubble.
We use the constants
\begin{equation}\label{eq:cosine-bubble}
  A = 0.5 \quad a = \SI{250}{\m} \quad x_c = \SI{500}{\m} \quad z_c = \SI{350}{\m}.
\end{equation}
We use a constant viscosity of $\mu = 0.1$.

\sidetitle{Two Bubbles}
The two bubbles scenario consists of a small cold air bubble that is on top of a large warm air bubble~\cite{robert1993bubble,muller2010adaptive}.
We use the same procedure as before to compute the initial conditions, again with $\backgroundPotT = 300$ but with a pertubation of the form

### 2019-01-21 writing/thesis/pages/abstract.tex
Small tweaks.
Include corrections from proof reader.

\chapter{\abstractname}
This thesis is concerned with the simulation of clouds with the ExaHyPE framework.
ExaHyPE is an engine for hyperbolic partial differential equations that uses the ADER-DG scheme.

We use the compressible Navier-Stokes equations for computational fluid dynamics.
Due to the diffusive components, the Navier-Stokes equations are not hyperbolic.
We thus need to modify parts of the numerical scheme.
Furthermore, we extend the equation set to the reactive compressible Navier-Stokes equations which include a source term that models chemical reactions.

We develop a gradient-based, global adaptive mesh refinement indicator that finds cells with unusually steep gradients.

Our implementation is evaluated for standard two and three dimensional fluid dynamics test cases, for flows over a realistic background atmosphere and for a reactive detonation wave.
Furthermore, we perform a convergence test and a time-to-solution benchmark that evaluates our proposed AMR criterion.

>>>>

\chapter{\abstractname}
This thesis is concerned with the simulation of clouds using the ExaHyPE framework.
ExaHyPE is an engine for hyperbolic partial differential equations that uses the ADER-DG scheme.

We use the compressible Navier-Stokes equations for computational fluid dynamics.
Due to the diffusive components, the Navier-Stokes equations are not hyperbolic.
We thus need to modify parts of the numerical scheme.
In addition, we extend the equation set to the reactive compressible Navier-Stokes equations which include a source term that models chemical reactions.

We develop a gradient-based, global adaptive mesh refinement (AMR) indicator that finds cells with unusually large gradients.

Our implementation is evaluated for standard two and three dimensional fluid dynamics test cases, for flows over a realistic background atmosphere and for a reactive detonation wave.
Additionally, we perform a convergence test and a time-to-solution benchmark that evaluates our proposed AMR criterion.
